\documentclass{article}
\usepackage[margin=0.6in]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\newcommand{\answeritem}{\global\answertrue\item}
\newcommand{\perhapsanswer}{%
  \ifanswer
    $\blacksquare$ \global\answerfalse
  \else
    $\square$ \global\answerfalse
  \fi
}
\newif\ifanswer

\begin{document}

  \title{605.601— Foundations of Software Engineering:Quiz 01}
  \author{Sabbir Ahmed}

  \maketitle

  \section*{Part 1} Answer all five questions below.

  \begin{enumerate}

    \item Choose the correct option in terms of the ‘code of ethics’ you know or recognize for issues related to professional responsibility of a software engineer
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Confidentiality
      \item Intellectual property rights
      \answeritem Both confidentiality \& intellectual property rights
      \item Managing Client Relationships
    \end{enumerate}

    \item Software Development Methodology includes a model and a framework
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \answeritem True
      \item False
    \end{enumerate}

    \item What is a Software?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Software is a set of programs
      \item Software is documentation and configuration of data
      \answeritem Software is set of programs, documentation \& configuration of data
      \item None of the mentioned
    \end{enumerate}

    \item Which one of these issues or points does not necessarily account for software failure?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Increasing demand
      \answeritem Low expectation
      \item Increasing supply
      \item Less reliable and expensive
    \end{enumerate}

    \item Which of the following is not software project management goal?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Keeping overall costs within budget
      \item Delivering the software to the customer at the agreed time
      \item Maintaining a happy and well-functioning development team
      \answeritem Avoiding customer complaints
    \end{enumerate}

  \end{enumerate}

  \section*{Part 2} Please use your own words to answer the following questions. Please make sure that your answer is clear, concise, and legible.

  \begin{enumerate}[resume]

    \item
    \begin{enumerate}
      \item What is a software development process?

      A software development process is a coherent set of activities grouped in distinct phases that lead to a software system's production.

      \item Why would you use it?

      A software development process is required to ensure that a software product is adequately implemented according to the specified requirements. The process also guarantees the proper maintenance of the software product after being shipped to the customers.

      \item What do you think are the common phases or stages of software development process?

      The common phases of a software development process are;
      \begin{enumerate}
        \item requirements gathering
        \item analysis and design
        \item implementation
        \item testing
        \item verification and acceptance
        \item maintenance
      \end{enumerate}

      \item Name and describe in a few sentences a software development process that is most commonly used in the industry today.

      A typical software development process in the industry is the agile methodology. Agile methodologies advocate for frequent collaboration between the development teams and their end-users through iterative development. Several derivations of the method exist to account for company or product constraints; however, the general principles of adaptive planning, evolutionary development, and continuous integration are maintained.

    \end{enumerate}
  \end{enumerate}

\end{document}
