\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{titling}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\newcommand{\answeritem}{\global\answertrue\item}
\newcommand{\perhapsanswer}{%
  \ifanswer
    $\blacksquare$ \global\answerfalse
  \else
    $\square$ \global\answerfalse
  \fi
}
\newif\ifanswer

\setstretch{1.5}
\setlength{\droptitle}{-8em}

\begin{document}

  \title{605.601 - Foundations of Software Engineering \\ Quiz 02\vspace{-0.5em}}
  \author{Sabbir Ahmed}
  \maketitle
  \vspace{-1em}

  \section*{Part 1} Answer any five questions below.

  \begin{enumerate}

    \item What kind of approach was introduced for elicitation and modelling to give a functional view of the system or software?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Object Oriented Design
      \answeritem Use Cases
      \item Fusion
      \item Object Modeling Technique
    \end{enumerate}

    \item A stakeholder is anyone who will purchase the completed software system under development.
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item True
      \answeritem False
    \end{enumerate}

    \item Which one of the following is not a step of requirement engineering?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \answeritem Elicitation
      \item Design
      \item Analysis
      \item Documentation
    \end{enumerate}

    \item Which model in software modeling using UML depicts the dynamic behavior of the software?
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item Context Model
      \item Object Model
      \item Data Model
      \item Sequence Model
    \end{enumerate}

    \item A Use-case actor is always a person having a role that different people may play.
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \item True
      \answeritem False
    \end{enumerate}

    \item Project management is less important for modern software development since most projects are successful and completed on time
    \begin{enumerate}[start=1,align=left,label={\protect\perhapsanswer(\alph*)}]
      \answeritem True
      \item False
    \end{enumerate}

  \end{enumerate}

  \section*{Part 2}

  \begin{enumerate}[resume]

    \item Please answer in your own words:

    \begin{enumerate}
      \item As a software engineer, do you have any responsibility or accountability to make sure that the software you and your team are developing – must be delivered on time and within budget? Please justify;

      \item If you are a part of an Agile development team, how much importance or attention would you like to offer in order for your team to gather and analyze the requirements? Please explain.
    \end{enumerate}

  \end{enumerate}


\end{document}
