\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{titling}

\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\newcommand{\answeritem}{\global\answertrue\item}
\newcommand{\perhapsanswer}{%
  \ifanswer
    $\blacksquare$ \global\answerfalse
  \else
    $\square$ \global\answerfalse
  \fi
}
\newif\ifanswer

\setstretch{1.5}
\setlength{\droptitle}{-8em}

\begin{document}

  \title{605.621 - Foundations of Algorithms \\ Assignment 01\vspace{-0.5em}}
  \author{Sabbir Ahmed}
  \maketitle
  \vspace{-1em}

  \section*{Part 1} Please answer all ten (10) questions below.

  \begin{enumerate}

    % 1
    \item {[10 pts, insertion sort bug]}

    Find the bug in the insertion code given in module 1 lecture notes. Show with an example where the bug causes the sort happening incorrectly.

    % 2
    \item {[10 pts, proof by induction]}

    Suppose we have the following sequence: $a_1=1, a_2=3, a_k=a_{k-2} + 2a_{k-1}$, for all integers $k\ge3$. Show that for all integers $n\ge1$, given the sequence $a_1, a_2, ..., a_k$ as defined above, $a_n$ is always odd.

    % 3
    \item {[10 pts, proof by contradiction]}

    Prove that $log_2(3)$ is irrational.

    % 4
    \item {[30 pts, time and space complexity]}
    
    For insertion sort algorithm,
    \begin{enumerate}
      \item Compute the best-case time complexity
      \item Compute the worst-case time complexity
      \item Compute the space complexity
    \end{enumerate}

    % 5
    \item {[40 pts, algorithm analysis]}

    The \textbf{loop invariance} concept is explained on page 19 of the textbook. During our programming of the insertion sort algorithm in class, we implicitly used this concept while we design/program the algorithm loops in a Python script.

    Read the relevant pages on the textbook and use loop invariance to solve problem 2-2 (page 40).

  \end{enumerate}

\end{document}
